A global sensitivity analysis of model uncertainty in aeroelastic wind turbine models

### 2020-02-20 Papers/Torque_Paper/content.tex
minor edits and accepting Prashant's changes to Torque article

\subsection{PCE-based Sobol indices computation}
The PCE approximation $f^{K}(\mathbf{z})$ of the computational model $Y$ can be defined as a weighted sum of multivariate polynomials in $\mathbf{z}$ \cite{RSmith}
\begin{equation}\label{eq:PCE}
Y= f(\mathbf{z}) \approx f^{K}(\mathbf{z}) = \sum_{|\mathbf{k}| = 0}^K w_{\mathbf{k}}\Psi_{\mathbf{k}}(\mathbf{z}),
\end{equation}
where $\mathbf{k}\in \mathbb{N}_0^M$ is a $M$-dimensional multi-index with magnitude $|\mathbf{k}| = k_1+k_2 + ... + k_M$ and $\Psi_{\mathbf{k}}(\mathbf{z})$ a multivariate polynomial basis, which is computed using a tensor product of univariate polynomials $\psi_{k_i}^{(i)}(z_i)$
\begin{equation}
\Psi_{\mathbf{k}}(\mathbf{z}) := \prod_{i=1}^M\psi_{k_i}^{(i)}(z_i),
\end{equation}
where $k_i$ is the order of the univariate polynomial. The choice of the univariate orthogonal polynomial $\psi_{k_i}$ depends on the type of random variable, $z_i$. For example, for uniform random variables the Legendre family of polynomials is used \cite{Xiu2002}. 
%The multivariate polynomials $\Psi_{\mathbf{k}}(\mathbf{z})$ is orthogonal with respect to the joint probability distribution of random vector $\mathbf{z}$. 
The total number of terms $K$ in \eqref{eq:PCE} can be based for example on a standard truncation scheme in which all polynomials in $M$ random variable with order less than or equal to $p$ are used. More advanced truncation schemes such as the maximum interaction or a hyperbolic scheme can also be employed depending on the application, see \cite{BlatmanThesis}. \todo{which one is used?}

@@

\subsection{NM80}
\subsubsection{Operating conditions}
As a reference wind turbine, we consider the 2MW NM80 turbine from the DANAERO project \cite{Troldborg2013} with a blade radius of 38.8 m. \todo{Add operating conditions}

@@

\subsubsection{Verification of sensitivity analysis method}
We perform a convergence study for both OLS and LARS in order to determine which method converges fastest with increasing number of uncertain parameters and number of model runs. For this purpose we only include uncertainty in the chord. The NURBS representation is defined by 9 control points, of which a subset is taken uncertain (see Fig.\ \ref{perturbed_samples}(a)). In Fig.\ \ref{convergence_dim}, we show for both OLS and LARS the convergence behaviour of the total Sobol indices of the power output with increasing number of model evaluations $N$ (runs of the AeroModule), and with increasing number of uncertain inputs $M$ (4, 6 and 8 control points). We see that both methods scale well with increasing dimensions, however, LARS exhibits faster convergence than OLS in high dimensions due to its adaptive nature, and will be the method of choice in the subsequent tests. \todo{what is a reasonable choice of $N$, depending on $M$?}

@@

\subsubsection{Sensitivity analysis results}
In Fig.~\ref{sensitivity_analysis}, we show the total order Sobol indices as a measure of the sensitivity of the different geometric and model parameters on the power output. In this case, we observe the important result that the uncertainty in model parameters is at least as important as the uncertainty in geometric parameters. Furthermore, within the geometric parameters, the chord distribution shows a significantly higher sensitivity compared to the twist variables

@@

\subsection{AVATAR}

>>>>

\subsection{PCE-based Sobol indices computation}
The PCE approximation $f^{K}(\mathbf{z})$ of the computational model $Y$ can be defined as a weighted sum of multivariate polynomials in $\mathbf{z}$ \cite{RSmith}
\begin{equation}\label{eq:PCE}
Y= f(\mathbf{z}) \approx f^{K}(\mathbf{z}) = \sum_{|\mathbf{k}| = 0}^K w_{\mathbf{k}}\Psi_{\mathbf{k}}(\mathbf{z}),
\end{equation}
where $\mathbf{k}\in \mathbb{N}_0^M$ is a $M$-dimensional multi-index with magnitude $|\mathbf{k}| = k_1+k_2 + ... + k_M$ and $\Psi_{\mathbf{k}}(\mathbf{z})$ a multivariate polynomial basis, which is computed using a tensor product of univariate polynomials $\psi_{k_i}^{(i)}(z_i)$
\begin{equation}
\Psi_{\mathbf{k}}(\mathbf{z}) := \prod_{i=1}^M\psi_{k_i}^{(i)}(z_i),
\end{equation}
where $k_i$ is the order of the univariate polynomial. The choice of the univariate orthogonal polynomial $\psi_{k_i}$ depends on the type of random variable, $z_i$. For example, for uniform random variables the Legendre family of polynomials is used \cite{Xiu2002}. 
%The multivariate polynomials $\Psi_{\mathbf{k}}(\mathbf{z})$ is orthogonal with respect to the joint probability distribution of random vector $\mathbf{z}$. 
The total number of terms $K$ in \eqref{eq:PCE} can be based for example on a standard truncation scheme in which all polynomials in $M$ random variable with order less than or equal to $p$ are used. In this work, a more advanced truncation schemes known as the hyperbolic truncation scheme is used \cite{BlatmanThesis}. 
%\textcolor{red}{In this work, we use the hyperbolic truncation scheme with a $qnorm = 0.75$.}

@@

\subsection{NM80}
\subsubsection{Operating conditions}
As a reference wind turbine, we consider the 2MW NM80 onshore turbine from the DANAERO project \cite{Troldborg2013} with a blade radius of 38.8 m and a hub height of 57.19 m with the mean wind speed at hub height set as 6.1m/s. Furthermore, the RPM is set at 12.3$/\text{min.}$, the blade pitch angle at 0.15 $\deg$ and the turbine yaw angle is zero.

@@

\subsubsection{Verification of sensitivity analysis method}
We perform a convergence study for both OLS and LARS in order to determine which method converges fastest with increasing number of uncertain parameters and number of model runs. For this purpose we only include uncertainty in the chord and set the power production as the quantity of interest. The NURBS representation is defined by 9 control points, of which a subset is taken uncertain (see Fig.\ \ref{perturbed_samples}(a)). In Fig.\ \ref{convergence_dim}, we show for both OLS and LARS the convergence behaviour of the total Sobol indices of the power output with increasing number of model evaluations $N$ (runs of the AeroModule), and with increasing number of uncertain inputs $M$ (4, 6 and 8 control points). We see that both methods scale well with increasing dimensions, however, LARS exhibits faster convergence than OLS in high dimensions due to its adaptive nature, and will be the method of choice in the subsequent tests. For both OLS and LARS, we start with a small $N$ and double it with every iteration until converged values of Sobol indices are obtained.
%}  \todo{what is a reasonable choice of $N$, depending on $M$?}

@@

\subsubsection{Sensitivity analysis results}
In Fig.~\ref{sensitivity_analysis}, we show the total order Sobol indices as a measure of the sensitivity of the different geometric and model parameters on the power output computed using LARS with $M=19$ uncertain parameter chosen from Fig.~\ref{perturbed_samples}. We found that $N=1024$ model evaluations were sufficient to achieve converged indices. In this case, we observe the important result that the uncertainty in model parameters is at least as important as the uncertainty in geometric parameters. Furthermore, within the geometric parameters, the chord distribution shows a significantly higher sensitivity compared to the twist variables.

@@
